Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-12 docs/proposals/NSF/ProjDesc/ProjDesc.tex
worked on transport

While the tortuosity characterizes the amount of winding of streamlines,
the dispersion characterizes the spreading of the trajectories.
Dispersion can be descried using the first- and second-ensemble moments
of the trajectories
\begin{align}
  \langle \lambda \rangle (t) = \frac{1}{N_p} \sum_{j=1}^{N_p}
    \lambda_j(t), \quad \sigma_\lambda^{2}(t) = \frac{1}{N_p}
    \sum_{j=1}^{N_p} (\lambda_j(t) - \langle \lambda \rangle(t))^2
\end{align}
Then, the standard deviation $\sigma_\lambda$ characterizes the
dispersion, and long-time behavior typically results in a power-law
scaling $\sigma_\lambda \sim t^{\alpha}$. The power $\alpha$, which
quantifies the long-time spreading, is $1$ in a ballistic regime which
occurs at early times when the trajectories have not explored much of
the porous geometry. However, once the trajectories have traversed a few
grains, dispersion in porous media is often super-dispersive with
$\alpha \in (0.5,1)$~\cite{}. In Figure~\ref{fig:100dispersion}, we
investigate spreading at several different porosities of a porous
geometry that is initialized with 100 grains.
\begin{wrapfigure}[]{r}{0.5\textwidth}
  \includegraphics[width=0.5\textwidth]{figs/100b_second_moment_long_ref}
  \caption{\label{fig:100dispersion} The particle spreading,
  characterized with the standard deviation $\sigma_\lambda(t)$ at 6
  different porosities. At early times, the line of slope 1 (black
  curve) indicates a ballistic motion. Then, at later times, the
  dispersion becomes super-dispersive at a rate indicated by the slope
  of the black dashed-dotted line.}
\end{wrapfigure}

The rate of dispersion is known to be directly related to the pore size
distribution~\cite{dea-qua-bir-jua2018}. Therefore, there is a need to
understand how erosion affects the pore size distribution. In
preliminary, we define pore sizes between grains that share an edge of a
Delanuay triangulation with vertices at the grain centers
(Figure~\ref{fig:triangluation}). In Figure~\ref{fig:poreStats}, we plot
the mean and variance of the pore sizes as a function of porosity.

>>>>

While the tortuosity characterizes the amount of winding of streamlines,
the dispersion characterizes the spreading of the trajectories.
Dispersion can be descried using the first- and second-ensemble moments
of the trajectories
\begin{align}
  \langle \lambda \rangle (t) = \frac{1}{N_p} \sum_{j=1}^{N_p}
    \lambda_j(t), \quad \sigma_\lambda^{2}(t) = \frac{1}{N_p}
    \sum_{j=1}^{N_p} (\lambda_j(t) - \langle \lambda \rangle(t))^2
\end{align}
Then, the standard deviation $\sigma_\lambda$ characterizes the
dispersion, and long-time behavior typically results in a power-law
scaling $\sigma_\lambda \sim t^{\alpha}$. The power $\alpha$, which
quantifies the long-time spreading, is $1$ in a ballistic regime which
occurs at early times when the trajectories have not explored much of
the porous geometry. However, once the trajectories have traversed a few
grains, dispersion in porous media is often super-dispersive with
$\alpha \in (0.5,1)$~\cite{}. In Figure~\ref{fig:100dispersion}, we
investigate spreading at several different porosities of a porous
geometry that is initialized with 100 grains.
\begin{wrapfigure}[]{r}{0.4\textwidth}
  \includegraphics[width=0.4\textwidth]{figs/100b_second_moment_long_ref}
  \caption{\label{fig:100dispersion} The particle spreading,
  characterized with the standard deviation $\sigma_\lambda(t)$ at 6
  different porosities. At early times, the line of slope 1 (black
  curve) indicates a ballistic motion. Then, at later times, the
  dispersion becomes super-dispersive at a rate indicated by the slope
  of the black dashed-dotted line.}
\end{wrapfigure}

The rate of dispersion is known to be directly related to the pore size
distribution~\cite{dea-qua-bir-jua2018}. Therefore, there is a need to
understand how erosion affects the pore size distribution. In
preliminary, we define pore sizes between grains that share an edge of a
Delanuay triangulation with vertices at the grain centers
(Figure~\ref{fig:triangluation}). In Figure~\ref{fig:poreStats}, we plot
the mean and variance of the pore sizes as a function of porosity. To
better understand how erosion affects the anomalous dispersion, we
propose to investigate how the pores grow with erosion. We will start
with the notion of a {\em porelet} which is an individual
Hagen-Poiseuille flow in each of the individual
pores~\cite{dea-qua-bib-jua2018}. By computing the strength of the
individual Hagen-Poiseuille flows, the shear rate can be computed, and
thus the rate of erosion can be computed.
